A dynamic programming algorithm to assess the reliability of inferences from network enrichment analysis

### 2018-05-03 doc/genesetdp.tex
text

Most GEA methods are investigating the overlap between the investigated set of analytes, the {\em query set}, and a functional module, the {\em pathway set}, using hypergeometric test or a Fisher's exact test. However, variants such as Gene Set Enrichment Analysis (GSEA)\cite{subramanian2005gene} also includes information on expression levels of the analytes of the query set.

@@

Since these methods of network randomization are not completely random, they introduce bias in the statistics that may or may not be desired, as the set of random networks is much larger than the ones of random networks that retain a certain topological characteristic. Also, the somewhat inaccessible process used to randomize networks produces statistics that are equaly hard to interpret as the null hypothesis becomes complex. FIXME: Should this be in the dicussion?

@@

As a representative example for we chose the ``Glycolysis/Gluconeogenesis'' pathway, as it contained a number of genes that coincided with the median number of genes of the pathway definition file.

@@

FIXME: Is it ok to state the method is more accurate if we are testing it agains a different null hypothesis

It should be noted that our outlined method a score distribution could work well with Monte Carlo-sampling as well, by sampling random gene sets of size $Q$, which would be faster and easier to understand than the theory of how to permute networks.

>>>>

Most GEA methods are assessing the overlap between the investigated set of analytes, the {\em query set}, and a functional module, the {\em pathway set}, using hypergeometric test or a Fisher's exact test. However, variants such as Gene Set Enrichment Analysis (GSEA)\cite{subramanian2005gene} also includes information on expression levels of the analytes of the query set.

@@

Since these methods of network randomization are not completely random, they introduce bias in the statistics that may or may not be desirable, as the set of random networks is much larger than the ones of random networks that retain a certain topological characteristic. Also, the somewhat inaccessible process used to randomize networks produces statistics that are equaly hard to interpret as the null hypothesis becomes complex. FIXME: Should this be in the dicussion?

@@

As a representative example for we chose the ``Glycolysis/Gluconeogenesis'' pathway, as it contained a number of genes that coincided with the median number of genes of the pathways in the definition file.

@@

FIXME: Is it ok to state the method is more accurate if we are testing it agains a different null hypothesis?

It should be noted that our outlined method a could work well with Monte Carlo-sampling, by sampling random gene sets of size $Q$, which would be faster and easier to understand than the theory of how to permute networks.
